---
title: "R Notebook"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

```{r}
library(tidyverse)
library(sjmisc)
library(ggplot2)
library(skimr)
library(kableExtra)
```

Exercise 3: Explore and Wrangle Data

Questions 1, 2, 3:

```{r}
f = "https://raw.githubusercontent.com/difiore/ada-2022-datasets/main/data-wrangling.csv"
d= read_csv(f, col_names = TRUE)
bsd = d$Body_mass_male_mean/d$Body_mass_female_mean
sex_ratio = d$AdultFemale/d$AdultMales
radius = sqrt(pi/d$HomeRange_km2)
diameter = 2*radius
di = d$DayLength_km/diameter

```

Question 4:Plot the relationship between day range length and time spent moving, for these primate species overall and by family.

```{r}
ggplot(data = d, aes(x = DayLength_km, y = Move))+
  geom_point(na.rm = TRUE) 

ggplot(data = d, aes(x = DayLength_km, y = Move))+
  geom_point(na.rm = TRUE) +
  facet_wrap(~Family)

```

Question 5: Plot the relationship between day range length and time group size, overall and by family. 

```{r}
ggplot(data = d, aes(x = DayLength_km, y = MeanGroupSize))+
  geom_point(na.rm = TRUE) 

ggplot(data = d, aes(x = DayLength_km, y = MeanGroupSize))+
  geom_point(na.rm = TRUE) +
  facet_wrap(~Family)
```

Question 6: Plot the relationship between body size dimorphism and canine size dimorphism overall and by family.

```{r}
ggplot(data = d, aes(x = bsd, y = Canine_Dimorphism))+
  geom_point(na.rm = TRUE) 

ggplot(data = d, aes(x = bsd, y = Canine_Dimorphism))+
  geom_point(na.rm = TRUE) +
  facet_wrap(~Family)

```


Question 7:

```{r}

d2 = mutate(d, diet_strategy = case_when(
  Fruit >= 50 ~ "frugivore",
  Leaves >= 50 ~ "folivore",
  Fruit < 50 & Leaves < 50 ~ "omnivore"
))

ggplot(d2 %>% filter(!is.na(diet_strategy)), aes(x=diet_strategy, y=MeanGroupSize))+
  geom_boxplot()

```

Question 8:
```{r}
d3 = d %>% 
  mutate("Binomial" =paste(Genus, Species, sep=" ")) %>% 
  select(Binomial, Family, Genus, Brain_Size_Species_Mean, Body_mass_male_mean, Body_mass_female_mean) %>% 
  group_by(Family) %>% 
  summarise(avg1 = mean(Brain_Size_Species_Mean, na.rm = TRUE),
            avg2 = mean(Body_mass_male_mean, na.rm = TRUE),
            na.rm = TRUE) %>% 
  arrange(avg1)
d3
```

